---
title: "Analysis"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r}
naRatioThreshold <- 0.9
```
In this dataset, we have two kinds of features/reponses: 1. common features (1-24) 2. product features (25-48)

```{r, message=FALSE, warning=FALSE}
options(warn=-1)
require(data.table, quietly = TRUE)
require(plyr, quietly = TRUE)
require(dplyr, quietly = TRUE)

featureDescription <- fread("../data/description.csv", verbose = FALSE)

#the reason that five features are read as character is that they are actually binary variables easy to convert to 0,1
colclasses <- list(
  integer = c("ncodpers", "ind_nuevo", "antiguedad", "age",  "ind_actividad_cliente"),
  numeric = c("renta"),
  factor = c("ind_empleado", "pais_residencia", "indrel_1mes", "tiprel_1mes",  "canal_entrada", "cod_prov", "nomprov", "segmento"),
  # Binary variables
  character = c("sexo", "indrel", "indresi", "indext", "indfall", "conyuemp"), 
  date = c("fecha_dato", "ult_fec_cli_1t", "fecha_alta"))

naStrings <- c("", " ", "NA")
trainSet <- fread("../data/train.csv", verbose = FALSE, na.strings = naStrings)
nTrain <- nrow(trainSet)
```
Here is a summary of the whole dataset. 
```{r, , message=FALSE, warning=FALSE}
testSet <- fread("../data/test.csv", header = TRUE, verbose = FALSE, na.strings = naStrings)
nTest <- nrow(testSet)
commonFeatureNames <- colnames(trainSet)[1:24]
productFeatureNames <- colnames(trainSet)[25:48]
testSet[,productFeatureNames] <- NA
wholeSet <- rbindlist(list(trainSet, testSet))
summary(wholeSet[,(commonFeatureNames),with=FALSE])
```

Note that 
- **indresi** : Residence index (S (Yes) or N (No) if the residence country is the same than the bank country) 
- **indext** : Foreigner index (S (Yes) or N (No) if the customer's birth country is different than the bank country) 

may need special attention here, which is not taken care of right now.

It is weird that **antiguedad** has value being -999999, which is supposed to be another way of reprensenting NAs. For safety purpose, we sets it to zero.

```{r}
wholeSet[antiguedad == -999999, antiguedad := 0] %>% invisible
```

Let us remove the non-product features with too many NA values. The features to remove are

```{r}
naRatios <- colMeans(is.na(wholeSet[, 1:24, with = FALSE]))
featureNamesWithTooManyNas <- names(which(naRatios > naRatioThreshold))
for (feature in featureNamesWithTooManyNas){
  cat(sprintf("feature %s is removed, with NA ratio being %.3f: %s\n", feature, naRatios[feature], featureDescription$Description[featureDescription$`Column Name` == feature]))
}
wholeSet[, (featureNamesWithTooManyNas) := NULL] %>% invisible
commonFeatureNames <- commonFeatureNames[!(commonFeatureNames %in% featureNamesWithTooManyNas)]
```

Then we remove the numeric features with very little variance:

```{r remove_col_with_little_var}
withLittleVariance <- sapply(wholeSet[,commonFeatureNames,with=FALSE], function(x) is.numeric(x) && var(x, na.rm = TRUE) < 0.01)
featureNamesWithLittleVar <- commonFeatureNames[which(withLittleVariance)]
for (feature in featureNamesWithLittleVar){
  cat(sprintf("feature %s is removed.\n", feature))  
}

wholeSet[, (featureNamesWithLittleVar) := NULL] %>% invisible
commonFeatureNames <- commonFeatureNames[!(commonFeatureNames %in% featureNamesWithLittleVar)]
```

Now let us look at the features with NAs
```{r check_na_ratio}
CheckFeaturesWithNas <- function(wholeSet, commonFeatureNames, productFeatureNames){
  nTotal <- wholeSet %>% nrow
  naRatioCommonFeatures <- sapply(wholeSet[, commonFeatureNames, with=FALSE], function(feature) mean(is.na(feature)))
  naRatioProductFeatures <- sapply(wholeSet[fecha_dato != "2016-06-28", productFeatureNames, with=FALSE], function(feature) mean(is.na(feature)))
  naRatio <- c(naRatioCommonFeatures, naRatioProductFeatures)

  hasNa <- naRatio != 0
  featureNamesWithNa <- colnames(wholeSet)[hasNa]
  naRatio <- naRatio[hasNa]
  if (length(naRatio) == 0){
    cat("No NAs!")
    return()
  }

  for(i in 1:length(naRatio)){
    featureName <- featureNamesWithNa[i]
    cat(sprintf("Col %s has %d NAs with ratio being %.3f %%: %s\n", featureName, ceiling(naRatio[i] * nrow(wholeSet)), naRatio[i] * 100, featureDescription$Description[featureDescription$`Column Name` == featureName]))
  }
}

CheckFeaturesWithNas(wholeSet, commonFeatureNames, productFeatureNames)
```

It looks very suspiciout that some features have the same number of NAs. It turns out that it is the same set of samples that have those missing features: `ind_empleado`, `pais_residencia`, `age`, `fecha_alta`, `ind_nuevo`, `antiguedad`, `indrel`, `indresi`, `indext`, `indfall`.

```{r}
areNas <- is.na(wholeSet$ind_empleado)
all(areNas == is.na(wholeSet$pais_residencia))
all(areNas == is.na(wholeSet$age))
all(areNas == is.na(wholeSet$fecha_alta))
all(areNas == is.na(wholeSet$ind_nuevo))
all(areNas == is.na(wholeSet$antiguedad))
all(areNas == is.na(wholeSet$indrel))
all(areNas == is.na(wholeSet$indresi))
all(areNas == is.na(wholeSet$indext))
all(areNas == is.na(wholeSet$indfall))
```

Now let us check these samples:

```{r}
naDataTable <- wholeSet[areNas, ]
sprintf("Total number of samples is %d", nrow(naDataTable)) %>% cat
tail(naDataTable[,1:24, with = FALSE], 5)
```

It looks like these samples misses all common features. Let us verify the hypo:

```{r}
allValuesMissing <- sapply(commonFeatureNames[3:length(commonFeatureNames)], function(featureName){
  all(is.na(naDataTable[[featureName]]))
})

all(allValuesMissing)
```

It seems that maybe all data associated with these ncodpers in this set are corrupted. Let us verify:

```{r}
ncodpers <- naDataTable$ncodpers %>% unique
cat(sprintf("We have %d unique customers", length(ncodpers)))
withIds <- wholeSet$ncodpers %in% ncodpers
ntotalTransactions <- withIds %>% sum

cat(sprintf("We have %d instances of this customers in the whole dataset, while %d of those instances miss all common features", ntotalTransactions, nrow(naDataTable)))
```

So this hypo is not entirely true, but almost true. 

```{r}
all(which(areNas) < nrow(trainSet))
!any(testSet$ncodpers, ncodpers)
```

As the results shows, all instances with all features missing are in train dataset, but in test set we have instances to predict associated with those customers, so it is not safe to remove all 29707, but safe to remove 27734 instances.

```{r}
dim(wholeSet)
wholeSet <- wholeSet[!areNas, ]
dim(wholeSet)
```

Now we check again which features still have missing values

```{r}
CheckFeaturesWithNas(wholeSet, commonFeatureNames, productFeatureNames)
```

Now we preprocess the dataset, together with the test set.

```{r}
wholeSet <- wholeSet[order(ncodpers, fecha_dato)]
LocalImpute <-function(values){
  if (length(values) == 1) return(values)
  for (i in 2:length(values)){
    if (is.na(values[i])) values[i] = values[i-1]
  }
  
  for (i in (length(values)-1):1){
    if (is.na(values[i])) values[i] = values[i+1]
  }
  return (values)
}

TopFreqItem <-function(values){
  return(names(which.max(table(values)))[1])
}

```

As we can see, we have 5 customers with gender info REALLY missing, we impute them with the mode number.

```{r sexo}
uids <- wholeSet$ncodpers[is.na(wholeSet$sexo)] %>% unique()
print(uids)
wholeSet$sexo[is.na(wholeSet$sexo)] <- "V"
table(wholeSet$sexo)
```


```{r}
# indrel_1mes: Customer type at the beginning of the month ,1 (First/Primary customer), 2 (co-owner ),P (Potential),3 (former primary), 4(former co-owner)
table(wholeSet$indrel_1mes, useNA = "always")
wholeSet[indrel_1mes != "P", indrel_1mes := sapply(indrel_1mes, substr, start = 1, stop = 1)] %>% invisible() # "X.0" => "X"
table(wholeSet$indrel_1mes, useNA = "always")
wholeSet[, indrel_1mes := LocalImpute(indrel_1mes), by=ncodpers] %>% invisible()
topFreqItem <- TopFreqItem(wholeSet$indrel_1mes)
wholeSet[is.na(indrel_1mes), indrel_1mes := topFreqItem] %>% invisible()
table(wholeSet$indrel_1mes, useNA = "always")
```

```{r}
# tiprel_1mes: Customer relation type at the beginning of the month, A (active), I (inactive), P (former customer),R (Potential)
table(wholeSet$tiprel_1mes, useNA = "always")
wholeSet[, tiprel_1mes := LocalImpute(tiprel_1mes), by=ncodpers] %>% invisible
table(wholeSet$tiprel_1mes, useNA = "always")
topFreqItem <- TopFreqItem(wholeSet$tiprel_1mes)
wholeSet[is.na(tiprel_1mes), tiprel_1mes := topFreqItem] %>% invisible
table(wholeSet$tiprel_1mes, useNA = "always")
```

```{r}
# canal_entrada: channel used by the customer to join
table(wholeSet$canal_entrada, useNA = "always")
wholeSet[, canal_entrada := LocalImpute(canal_entrada), by=ncodpers] %>% invisible
table(wholeSet$canal_entrada, useNA = "always")
topFreqItem <- TopFreqItem(wholeSet$canal_entrada)
wholeSet[is.na(canal_entrada), canal_entrada := topFreqItem] %>% invisible
table(wholeSet$canal_entrada, useNA = "always")
```

```{r}
# nomprov
table(wholeSet$nomprov, useNA = "always")
wholeSet[nomprov %>% is.na, nomprov := "UNKNOWN"] %>% invisible
wholeSet[, nomprov:= as.factor(nomprov)] %>% invisible
levels(wholeSet$nomprov) <- gsub(",", "", levels(wholeSet$nomprov))
table(wholeSet$nomprov, useNA = "always")

# cod_prov is redundant with nomprov
wholeSet <- wholeSet[, cod_prov := NULL]
commonFeatureNames <- commonFeatureNames[commonFeatureNames != "cod_prov"]
```


```{r}
summary(wholeSet$renta)
wholeSet[, renta := LocalImpute(renta), by=ncodpers] %>% invisible
summary(wholeSet$renta)
# renta 20.476 % missing is a very high number
wholeSet[renta %>% is.na, renta := median(wholeSet$renta, na.rm = TRUE)] %>% invisible
summary(wholeSet$renta)
```

```{r segmento}
print(featureDescription$Description[featureDescription$`Column Name` == "segmento"])
table(wholeSet$segmento, useNA = "always")
wholeSet[, segmento := LocalImpute(segmento), by=ncodpers] %>% invisible
table(wholeSet$segmento, useNA = "always")
topFreqItem <- TopFreqItem(wholeSet$segmento)
wholeSet[is.na(segmento), segmento := topFreqItem] %>% invisible
table(wholeSet$segmento, useNA = "always")
```


```{r}
# ind_nomina_ult1: product feature
wholeSet[, ind_nomina_ult1:=LocalImpute(ind_nomina_ult1)] %>% invisible
wholeSet[ind_nomina_ult1 %>% is.na, ind_nomina_ult1 := median(wholeSet$ind_nomina_ult1, na.rm = TRUE)] %>% invisible
wholeSet[fecha_dato == "2016-06-28", ind_nomina_ult1 := NA] %>% invisible
# ind_nom_pens_ult1: product feature
wholeSet[, ind_nom_pens_ult1 := LocalImpute(ind_nom_pens_ult1)] %>% invisible
wholeSet[ind_nom_pens_ult1 %>% is.na, ind_nom_pens_ult1 := median(wholeSet$ind_nom_pens_ult1, na.rm = TRUE)] %>% invisible
wholeSet[fecha_dato == "2016-06-28",ind_nom_pens_ult1 := NA] %>% invisible
```

Now we check if there still any missing values existing:

```{r}
CheckFeaturesWithNas(wholeSet, commonFeatureNames, productFeatureNames)
```

Now let us write the preprocessed datasets in files

```{r}
fwrite(wholeSet[fecha_dato != "2016-06-28"], "../data/preprocessed_train.csv", row.names = FALSE, quote = FALSE)
fwrite(wholeSet[fecha_dato == "2016-06-28", commonFeatureNames, with = FALSE], "../data/preprocessed_test.csv", row.names = FALSE, quote = FALSE)
```

